---
title: "OCTA Repeatability & Reliability — Clean Data"
author: "Jae Lee, Bruce Burkemper, Dom Grisafe, Grace Richter"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
csl: ophthalmology.csl
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(readxl)           # import excel files
library(tidyverse)
library(Hmisc)
library(ICC)              # inter class correlation
library(labelled)         # add variable labels 
library(kableExtra)       # HTML tables
library(tableone)         # table 1
library(cowplot)          # makes clean plots, with no distractions
theme_set(theme_cowplot())
library(gridExtra)        # allows grid arangement of dissimilar plots together
library(grid)             # add labels to top of grid.arrange plots
library(lme4)
library(citr)
```

#Data

This data was taken from a clean dataset that Jae emailed out on October 25th, 2019.

Jae organized the retina by six different methods. He created OCTA variable scores for the overall image, 2 hemifields, 4 quadrants, 6 sectors, 12 clock hours, and an annular ring.  

For now, only the overall image scoring will be analyzed.

I manually removed underscores from the names of each patient because it was causing uneven splitting of variables. This modified dataset is saved as all_quantification_10_24_19_DG.xlsx. I updated the names for every tab of the file.

```{r}
# function that formats variables appropriately
form_vars <- function(df){

  df %>% 
    
    select(NameSIDSexEyeRegion, everything()) %>% 
    
    mutate(
      
      # match first text (upper or lower case) of any length, followed by any underscore, all repeated any number of times
      name = str_extract(string = NameSIDSexEyeRegion, pattern = regex("^[[A-Za-z]*_]+")) %>% str_sub(end = -2),
      
      # matches everything starting with the first number
      SIDSexEyeRegion = str_extract(string = NameSIDSexEyeRegion, pattern = regex("[[0-9]*].*$"))
      
      ) %>% 
    
    # separate the remaining variables
    separate(
      col = SIDSexEyeRegion, sep = "_", remove = TRUE,
      into = c("ptid", "sex", "image_area", "image_method", "image_size", "exam_date", "exam_time", "eye_imaged")
      ) %>% 
    
    # filter blank rows
    filter(!is.na(exam_date)) %>% 
    
    # format variables appropriately
    mutate(
      
    time_stamp = paste(exam_date, exam_time) %>% as.POSIXct(format = "%m%d%Y %H%M%S"),

    exam_date = exam_date %>% as.Date(format = "%m%d%Y"),
    
    eye_imaged = case_when(
      eye_imaged == "OS" ~ 1,
      eye_imaged == "OD" ~ 2
    ) %>% factor(levels = c(1, 2), labels = c("Left", "Right")),
    
    female = case_when(
      sex == "F" ~ 1,
      sex == "M" ~ 2
    ) %>% factor(levels = c(1,2), labels = c("Female", "Male")),
    
    image_area = factor(image_area),

    image_method = factor(image_method),
    
    image_size = factor(image_size)
    
    ) %>% 
    
    # remove unnecessary variables
    select(-NameSIDSexEyeRegion, -sex, -exam_time) %>% 
    
    # arrange so in appropriate order
    arrange(ptid, eye_imaged, time_stamp) %>% 
    
    # create a measurement ID
    mutate(eyeid = group_indices(., ptid, eye_imaged)) %>% 
 
    # deidentify names
    select(-name) %>%
    
    # reorder variables appropriately
    select(ptid, female, eyeid, eye_imaged, image_area, image_size, image_method, exam_date, time_stamp, everything())
  
}
```


```{r, message = FALSE, warning = FALSE}
# create a function that removes the middle initial and titles from the names

setwd("/Volumes/GoogleDrive/My Drive/2017-2020 USC Roski Eye Institute/OCTA Repeatability - Grace Richter/Data/Jae")

data_octa_global <- read_xlsx("all_quantification_11_20_19.xlsx", sheet = "Global") %>% form_vars()
data_octa_annulus <- read_xlsx("all_quantification_11_20_19.xlsx", sheet = "Annulus") %>% form_vars()
data_octa_hemifield <- read_xlsx("all_quantification_11_20_19.xlsx", sheet = "Hemifield") %>% form_vars()
data_octa_quadrant <- read_xlsx("all_quantification_11_20_19.xlsx", sheet = "Quadrants") %>% form_vars()
data_octa_sector <- read_xlsx("all_quantification_11_20_19.xlsx", sheet = "Sector") %>% form_vars()
data_octa_clockhour <- read_xlsx("all_quantification_11_20_19.xlsx", sheet = "Clockhours") %>% form_vars()

data_octa_commercial <- read_xlsx("all_quantification_11_20_19.xlsx", sheet = "Commercial measurement") %>% form_vars()

# data_octa_commercial %>% glimpse
```


##Data Dictionary

```{r}

var_labels <- c(
  ptid = "Patient Identification",
  eyeid = "Eye Identification",
  eye_imaged = "Eye Imaged",
  time_stamp = "Exam Date and Time",
  name_last = "Last Name",
  name_first = "First Name",
  female = "Female Sex",
  image_area = "Image Area",
  image_method = "Image Method",
  image_size = "Image Size",
  v_diameter = "Vessel Diameter (units?)",
  v_area_density = "Vessel Area Density (units?)",
  v_skeleton_density = "Vessel Skeleton Density (units?)",
  v_perimeter_index = "Vessel Perimeter Index (units?)",
  v_complexity_index = "Vessel Complexity Index (units?)",
  flow_impair_zone = "Flow Impairment Zone (units?)",
  flux = "Flux (units?)"
)

# assign variable labels
data_dict <-  data.frame(var_labels) %>% 
  rownames_to_column("Variable")

# print html table
data_dict %>% 
  kable(col.names = c("Variable", "Label")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))
```


##Data Wrangling

###Rename Variables

This section will rename the variables to match the dictionary

```{r}

# function to rename OCTA variables
rename_octa_vars <- function(df){
  
  df %>% 
    
    # rename variables in existing format
    rename(
      v_diameter = starts_with("Vessel Diameter"),
      v_area_density = starts_with("Vessel Area Density"),
      v_skeleton_density = starts_with("Vessel Skeleton Density"),
      v_perimeter_index = starts_with("Vessel Perimeter Index"),
      v_complexity_index = starts_with("Vessel Complexity Index"),
      flow_impair_zone = starts_with("Flow Impiarment Zone"),
      flux = starts_with("Flux")
    ) 
 
}

data_octa_global_rename <- rename_octa_vars(data_octa_global)
data_octa_annulus_rename <- rename_octa_vars(data_octa_annulus)
# data_octa_hemifield_rename <- rename_octa_vars(data_octa_hemifield)
# data_octa_quadrant_rename <- rename_octa_vars(data_octa_quadrant)
# data_octa_sector_rename <- rename_octa_vars(data_octa_sector)
# data_octa_clockhour_rename <- rename_octa_vars(data_octa_clockhour)

# check out the renamed datasets
# data_octa_global_rename %>% glimpse
# data_octa_annulus_rename %>% glimpse

# commercial data had different variable names than the remaining variables
data_octa_commercial_rename <- data_octa_commercial %>% 
  # remove redundant commercial variables
  select(-`Eye Imaged`, -EyeRegion, -SID, -Sex, -`Time stamp`, -`Date Studied`) %>% 
  # rename the OCTA variables
  rename(
    
  ) 

# data_octa_commercial_rename %>% glimpse
```


###Label Variables

This section will label variables appropriately.

```{r}
# match variable labels with variable names in dataframe
label(data_octa_global_rename) <- as.list(var_labels[match(names(data_octa_global_rename), names(var_labels))])
label(data_octa_annulus_rename) <- as.list(var_labels[match(names(data_octa_annulus_rename), names(var_labels))])
# label(data_octa_hemifield_rename) <- as.list(var_labels[match(names(data_octa_hemifield_rename), names(var_labels))])
# label(data_octa_quadrant_rename) <- as.list(var_labels[match(names(data_octa_quadrant_rename), names(var_labels))])
# label(data_octa_sector_rename) <- as.list(var_labels[match(names(data_octa_sector_rename), names(var_labels))])
# label(data_octa_clockhour_rename) <- as.list(var_labels[match(names(data_octa_clockhour_rename), names(var_labels))])

label(data_octa_commercial_rename) <- as.list(var_labels[match(names(data_octa_commercial_rename), names(var_labels))])
```

#Single Eye Participation

Venugopal (2019) et al. ran the analysis including all eyes.[@venugopalRepeatabilityComparabilityPeripapillary2019] Then they re-ran the analysis after selecting only a single eye randomly from each individual. They conclude minimal difference in the findings, and they report only the analysis that includes all eyes: 

"The entire analysis was repeated considering one eye per subject (one eye was randomly chosen from subjects contributing both eyes for the primary analysis) and all the results were similar to the primary analysis."[@venugopalRepeatabilityComparabilityPeripapillary2019]

Based on this, analysis was performed on all eyes. If time permits, remove a random eye later.

```{r}
# function that gets just one eye
get_1eye <- function(df){
  
  # set the seed so working with a consistent dataset when randomly sampling left or right eye
  set.seed(123456)
  
  # select random 1 eye per person
  select_1eye <- df %>% 
    # get 1 eyeid
    group_by(ptid, eyeid) %>% sample_n(1) %>% select(ptid, eyeid) %>% ungroup() %>% 
    # sample 1 eye per individual
    group_by(ptid) %>% sample_n(1) %>% ungroup() %>% select(eyeid)
  
  # subset just the eyes in the selection above
  df %>% 
    left_join(x = select_1eye, by = "eyeid")
  
}

# get_1eye(data_octa_global_rename) %>% glimpse
```

#Subsetting Cohorts

##Intra-Visit Data

Intra-visit data was restricted to all measurements of the same eye that occurred on the same day. Some people had multiple measurements on multiple days. Their first day was included only.

```{r}
  
# Intra-Visit Data: create a dataset that selects only repeat measurements per day
data_octa_global_intra <- data_octa_global_rename %>% 
  
  # get 1 eye per person
  get_1eye() %>% 
  
  # create variable that indicates when there is a repeat measurement on the same day for each eye
  add_count(ptid, eye_imaged, exam_date, name = "n_msr_day") %>%
  filter(n_msr_day > 1) %>%
  
  # now that all the days with single measurements have been excluded, pick the first date with multiple measurements
  group_by(ptid, eye_imaged) %>%
  filter(exam_date == min(exam_date)) %>%
  # create an indicator variable that shows the order of measurements
  mutate(order_obs = seq(n())) %>% 
  ungroup() %>% 
  
  # create an indicator variable that shows the order of measurements
  group_by(ptid, eye_imaged) %>% 
  mutate(order_obs = seq(n())) %>% 
  ungroup()
```

##Inter-Visit Data

Inter-visit data was restricted to the first measurement of the day for each eye. If a patient had their right eye measured:    

- 3 times on Monday  

- 5 times on Wednesday, and  

- 2 times on Friday,  

Then they would have three observations in this dataset:

- The first measurement from Monday,  

- The first measurement from Wednesday, and  

- The first measurement from Friday.

```{r}
# Inter-Visit Data: create a dataset that selects the first measurement per day for each eye; there are no repeat measurements per day
data_octa_global_inter <- data_octa_global_rename %>% 
  
  # get 1 eye per person
  get_1eye() %>% 

  group_by(ptid, eye_imaged, exam_date) %>% 
  top_n(n = -1, wt = time_stamp) %>% 
  ungroup() %>% 
  # create an indicator variable that shows the order of measurements
  group_by(ptid, eye_imaged) %>% 
  mutate(order_obs = seq(n())) %>% 
  ungroup()
```

##Data Flow Chart

This tallies how many participants, how many eyes, and how many measurements are included in the two datasets.

```{r}

data.frame(
  
  dataset = c(
    "Intra-Visit",
    "Inter-Visit"
    ),
  
  #number of participants
  n_participants = c(
    data_octa_global_intra %>% group_by(ptid) %>% summarise(n()) %>% nrow(),
    data_octa_global_inter %>% group_by(ptid) %>% summarise(n()) %>% nrow()
    ),
  
  #number of eyes
  n_eyes = c(
    data_octa_global_intra %>% group_by(ptid, eye_imaged) %>% summarise(n()) %>% nrow(),
    data_octa_global_inter %>% group_by(ptid, eye_imaged) %>% summarise(n()) %>% nrow()
    ),
  
  # number of measurements
  n_measurements = c(
    data_octa_global_intra %>% nrow(),
    data_octa_global_inter %>% nrow()
    )
  
  ) %>%
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```


#Table 1: Summary Statistics

The Venugopal papers dichotomize glaucoma. But there's no glaucoma indicator variable in this data.

```{r}
print_table1 <- function(df){
  
  # variables to be included
  Vars_table1 <- c(
    "female", "eye_imaged", "image_area", "image_method", "image_size", 
    "v_diameter", "v_area_density", "v_skeleton_density", "v_perimeter_index", "v_complexity_index",
    "flow_impair_zone", "flux" 
  )
  # variables not included
  # "eyeid", "ptid", "name_last", "name_first", "time_stamp", "exam_date"
  
  # factor variables
  catVars_table1 <- c("female", "eye_imaged", "image_area", "image_method", "image_size")
  
  # exclude the following from table1: -SID, -Target, -Judge, -Date, -Meds, -Sx, -GLC
  df_table1 <- df %>% select(Vars_table1) %>% mutate(
    v_diameter = as.double(v_diameter),
    v_area_density = as.double(v_area_density),
    v_skeleton_density = as.double(v_skeleton_density),
    v_perimeter_index = as.double(v_perimeter_index),
    v_complexity_index = as.double(v_complexity_index),
    flow_impair_zone = as.double(flow_impair_zone),
    flux = as.double(flux)
  )
  
    ## export table 1 to html
    CreateTableOne(data = df_table1, vars = Vars_table1, factorVars = catVars_table1) %>% 
      print(exact = "stage", quote = FALSE, noSpaces = F, printToggle = F, varLabels = T) %>%
      kableone(digits = 3) %>%
      kable_styling(
        #bootstrap table classes
        bootstrap_options = c("striped", "hover")
      ) %>% 
      pack_rows("Sociodemographics", 2, 2) %>%
      pack_rows("Image Specifications", 3, 6) %>%
      pack_rows("OCTA Measures", 7, 13)
    
}
```

##Intra-Visit OCTA Summary Statistics

Variables in intra-visit dataset

```{r}
print_table1(data_octa_global_intra)
```

##Inter-Visit OCTA Summary Statistics

Variables in inter-visit dataset

```{r}
print_table1(data_octa_global_inter)
```


#Exploratory Data Analysis

##Scatter Plots{.tabset}

Scatter plots of OCTA variables by Subject ID for about 10 individuals randomly sampled. These are separated by intra- and inter-visit datasets, and are useful for understanding the data better.

```{r}

scat_octa_sid <- function(octa_var){

  # select a random sample of 10 SID's
  set.seed(1998)
  ptid_sample10 <- data_octa_global_rename %>% dplyr::filter(eye_imaged == "Right") %>% select(ptid) %>% sample_n(10) %>% unlist() %>% as.character()
  
grid.arrange(ncol = 2, left = var_labels[octa_var], top = "OCTA Measures by Patient", bottom = var_labels["ptid"],

             data_octa_global_intra %>%
               dplyr::filter((ptid %in% ptid_sample10) & (eye_imaged == "Right")) %>%
               ggplot(aes_string(x = "ptid", y = octa_var, group = "ptid")) +
               geom_point(alpha = 0.1, size = 5, color = "red") +
               ylim(min(data_octa_global_rename[[octa_var]]), max(data_octa_global_rename[[octa_var]])) +
               ggtitle("Intra") +
               ylab("") +
               xlab("") +
               theme(
                 legend.position = "none",
                 plot.title = element_text(hjust = 0.5),
                 axis.text.x = element_text(angle = 30, hjust = 1, size = 7)
                 ),
             
             data_octa_global_inter %>%
               dplyr::filter((ptid %in% ptid_sample10) & (eye_imaged == "Right")) %>%
               ggplot(aes_string(x = "ptid", y = octa_var, group = "ptid")) +
               geom_point(alpha = 0.1, size = 5, color = "blue") +
               ylim(min(data_octa_global_rename[[octa_var]]), max(data_octa_global_rename[[octa_var]])) +
               ggtitle("Inter") +
               ylab("") +
               xlab("") +
               theme(
                 legend.position = "none",
                 plot.title = element_text(hjust = 0.5),
                 axis.text.x = element_text(angle = 30, hjust = 1, size = 7)
                 )
)
}


# scat_octa_sid("v_area_density")

# data_octa_global_rename %>% glimpse
```

###Vessel Area Density
```{r}
scat_octa_sid("v_area_density")
```

###Vessel Diameter
```{r}
scat_octa_sid("v_diameter")
```

###Vessel Skeleton Density
```{r}
scat_octa_sid("v_skeleton_density")
```

###Vessel Perimeter Index
```{r}
scat_octa_sid("v_perimeter_index")
```

###Vessel Complexity Index
```{r}
scat_octa_sid("v_complexity_index")
```

###Flow Impairment Zone
```{r}
scat_octa_sid("flow_impair_zone")
```

###Flux
```{r}
scat_octa_sid("flux")
```

##Spaghetti Plots{.tabset}

Spaghetti plots of OCTA variables for about 10 individuals randomly sampled.

```{r}

spaghet_plot <- function(octa_var){

  # select a random sample of 10 patients
  set.seed(2448)
  ptid_sample10 <- data_octa_global_rename %>% dplyr::filter(eye_imaged == "Right") %>% select(ptid) %>% sample_n(10) %>% unlist() %>% as.character()

grid.arrange(ncol = 2, left = var_labels[octa_var], top = "OCTA Measures Over Time", bottom = var_labels["time_stamp"],
             
  data_octa_global_intra %>%
    dplyr::filter((ptid %in% ptid_sample10) & (eye_imaged == "Right")) %>%
    ggplot(aes_string(x = "time_stamp", y = octa_var, group = "ptid")) +
    geom_line(alpha = 0.5, size = 1, color = "red") +
    theme(legend.position = "none") +
    ylim(min(data_octa_global_rename[[octa_var]]), max(data_octa_global_rename[[octa_var]])) +
    ggtitle("Intra") +
    ylab("") +
    xlab("") +
    theme(
      legend.position = "none",
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(angle = 30, hjust = 1, size = 7)
    ),
  
  data_octa_global_inter %>%
    dplyr::filter((ptid %in% ptid_sample10) & (eye_imaged == "Right")) %>%
    ggplot(aes_string(x = "time_stamp", y = octa_var, group = "ptid")) +
    geom_line(alpha = 0.5, size = 1, color = "blue") +
    theme(legend.position = "none") +
    ylim(min(data_octa_global_rename[[octa_var]]), max(data_octa_global_rename[[octa_var]])) +
    ggtitle("Inter") +
    ylab("") +
    xlab("") +
    theme(
      legend.position = "none",
      plot.title = element_text(hjust = 0.5),
      axis.text.x = element_text(angle = 30, hjust = 1, size = 7)
    )
)
}

# spaghet_plot("v_area_density")
```

###Vessel Area Density
```{r}
spaghet_plot("v_area_density")
```

###Vessel Diameter
```{r}
spaghet_plot("v_diameter")
```

###Vessel Skeleton Density
```{r}
spaghet_plot("v_skeleton_density")
```

###Vessel Perimeter Index
```{r}
spaghet_plot("v_perimeter_index")
```

###Vessel Complexity Index
```{r}
spaghet_plot("v_complexity_index")
```

###Flow Impairment Zone
```{r}
spaghet_plot("flow_impair_zone")
```

###Flux
```{r}
spaghet_plot("flux")
```


#Bland-Altman Plots

Coding of Bland-Altman (BA) plots is explained in this [blog post](https://www.r-bloggers.com/bland-altmantukey-mean-difference-plots-using-ggplot2/).


##Intra-Visit Bland-Altman Plots{.tabset}

Intra-Visit Bland-Altman Plots of OCTA Variables


```{r, warning = FALSE}

# function that gets repeatability statistics
get_blandaltman <- function(df, octa_var){

  # make OCTA variable name
  octa_name <- var_labels[as.character(octa_var)]

  # intra-visit dataset
  data_blandalt <- df %>%

    # select only the measurement ID and the OCTA variable of interest
    select(ptid, eye_imaged, order_obs, eyeid, !!octa_var) %>%
        
    # only select the first 2 measurements
    filter(order_obs <= 2) %>% 

    group_by(ptid, eye_imaged) %>% 

    # calculate the difference and mean of measurements for each individual eye
    mutate(
      diff = diff(!!octa_var),
      mean = mean(!!octa_var)
      ) %>% ungroup() %>% 
    select(!!octa_var, everything())

  p_blandalt <- data_blandalt %>%
    ggplot(aes(x = mean, y = diff)) +
    scale_y_continuous(limits = c(-max(data_blandalt$diff)*1.5, max(data_blandalt$diff)*1.5)) +
    geom_point(size = 0.6, alpha = 0.9) +
    geom_hline(yintercept = mean(data_blandalt$diff), colour = "blue", size = 0.5, linetype = 2) +
    geom_hline(yintercept = mean(data_blandalt$diff) - (1.96 * sd(data_blandalt$diff)), colour = "red", size = 0.5, linetype = 2) +
    geom_hline(yintercept = mean(data_blandalt$diff) + (1.96 * sd(data_blandalt$diff)), colour = "red", size = 0.5, linetype = 2) +
    geom_text(aes(label = ifelse(
      diff >  mean(data_blandalt$diff) + (1.96 * sd(data_blandalt$diff)) |
      diff <  mean(data_blandalt$diff) - (1.96 * sd(data_blandalt$diff)),
        yes = as.character(eyeid), no = ''
      )), hjust = 0, vjust = 0, color = "darkgray", size = 3) +
    ggtitle("") +
    ylab("Difference in Measurements") +
    xlab("Average Measurements")

  p_diffdist <- data_blandalt %>%
    ggplot(aes(x = diff)) +
    geom_vline(xintercept = mean(data_blandalt$diff), colour = "blue", size = 0.5, linetype = 2) +
    geom_vline(xintercept = mean(data_blandalt$diff) - (1.96 * sd(data_blandalt$diff)), colour = "red", size = 0.5, linetype = 2) +
    geom_vline(xintercept = mean(data_blandalt$diff) + (1.96 * sd(data_blandalt$diff)), colour = "red", size = 0.5, linetype = 2) +
    geom_histogram(bins = 30, alpha = 0.9) +
    geom_text(aes(label = ifelse(
      diff >  mean(data_blandalt$diff) + (1.96 * sd(data_blandalt$diff)) |
      diff <  mean(data_blandalt$diff) - (1.96 * sd(data_blandalt$diff)),
        yes = as.character(eyeid), no = ''
      ), y = 6), hjust = 0, vjust = 0, color = "darkgray", size = 3) +
    scale_y_continuous(limits = c(0, 200), breaks = seq(0,200,50)) +
    scale_x_continuous(limits = c(-max(data_blandalt$diff)*1.5, max(data_blandalt$diff)*1.5)) +
    ggtitle("") +
    xlab("Difference in Measurements")

  grid.arrange(ncol = 2, top = textGrob(octa_name), p_blandalt, p_diffdist)
}

# get_blandaltman(df = data_octa_global_intra, octa_var = quo(v_area_density))
```


###Vessel Area Density
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_intra, octa_var = quo(v_area_density))
```

###Vessel Diameter
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_intra, octa_var = quo(v_diameter))
```

###Vessel Skeleton Density
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_intra, octa_var = quo(v_skeleton_density))
```

###Vessel Perimeter Index
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_intra, octa_var = quo(v_perimeter_index))
```

###Vessel Complexity Index
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_intra, octa_var = quo(v_complexity_index))
```

###Flow Impairment Zone
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_intra, octa_var = quo(flow_impair_zone))
```

###Flux
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_intra, octa_var = quo(flux))
```

###Outliers

Assess some of the observations outside of expected limits.

```{r}
data_octa_global_intra %>% 
  filter(eyeid %in% c(50, 51, 55, 56, 121, 122, 139, 140, 175, 176, 592, 593)) %>% 
  kable(digits = 3) %>%
  kable_styling(
    #bootstrap table classes
    bootstrap_options = c( "hover")
  )
```


##Inter-Visit Bland-Altman Plots{.tabset}

Inter-Visit Bland-Altman Plots of OCTA Variables

###Vessel Area Density
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_inter, octa_var = quo(v_area_density))
```

###Vessel Diameter
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_inter, octa_var = quo(v_diameter))
```

###Vessel Skeleton Density
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_inter, octa_var = quo(v_skeleton_density))
```

###Vessel Perimeter Index
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_inter, octa_var = quo(v_perimeter_index))
```

###Vessel Complexity Index
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_inter, octa_var = quo(v_complexity_index))
```

###Flow Impairment Zone
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_inter, octa_var = quo(flow_impair_zone))
```

###Flux
```{r, warning = FALSE}
get_blandaltman(df = data_octa_global_inter, octa_var = quo(flux))
```


###Outliers

Assess some of the observations outside of expected limits.

```{r}
data_octa_global_inter %>% 
  filter(eyeid %in% c(30, 32, 42, 45, 73, 75, 138, 141, 187, 188, 271, 272, 294, 295, 330, 331, 538, 535)) %>% 
  kable(digits = 3) %>%
  kable_styling(
    #bootstrap table classes
    bootstrap_options = c( "hover")
  )
```

#Intra-Visit Repeatability

Intra-Visit Repeatability of OCTA Variables

Calculate the within-subject standard deviation $(S_w)$, the within-subject coefficient of repeatability $(CR_w)$, and the within-subject coefficient of variation $(CV_w)$ for a given measurement variable $x$.  

```{r, include = FALSE}

# I tried to calculate the Sw manually and by using the one-way anova table, as instructed in Bland (1996). However, the two methods are producing different answers for Sw. Therefore, try to use the data in the paper to recalculate Sw for the sample data provided.

# after rechecking, it turns out my calculations were correct, but it's just the rounding errors combined with squaring and square-rooting to go from standard deviation to variance that caused the differences.

# data copied from Bland (1996)
df_bland96 <- matrix(c(190, 220, 200, 200, 202.50, 12.58, 220, 200, 240, 230, 222.50, 17.08, 260, 260, 240, 280, 260.00, 16.33, 210, 300, 280, 265, 263.75, 38.60, 270, 265, 280, 270, 271.25, 6.29, 280, 280, 270, 275, 276.25, 4.79, 260, 280, 280, 300, 280.00, 16.33, 275, 275, 275, 305, 282.50, 15.00, 280, 290, 300, 290, 290.00, 8.16, 320, 290, 300, 290, 300.00, 14.14, 300, 300, 310, 300, 302.50, 5.00, 270, 250, 330, 370, 305.00, 55.08, 320, 330, 330, 330, 327.50, 5.00, 335, 320, 335, 375, 341.25, 23.58, 350, 320, 340, 365, 343.75, 18.87, 360, 320, 350, 345, 343.75, 17.02, 330, 340, 380, 390, 360.00, 29.44, 335, 385, 360, 370, 362.50, 21.02, 400, 420, 425, 420, 416.25, 11.09, 430, 460, 480, 470, 460.00, 21.60), ncol = 6, byrow = T, dimnames = list(c(), c("1st", "2nd", "3rd", "4th", "Mean", "SD"))) %>%
  # convert to dataframe
  as.data.frame() %>%
  # make a column of children ID
  mutate(children = seq(1:20)) %>% select(children, everything())

# calculate the values manually
df_bland96 %>%
  # take the average of mean measurements and the average of the variances
  summarise(
    Xw = mean(Mean),
    Sw = sqrt(mean(SD^2))
    ) %>%
  # square root the variances
  mutate(
    CRw = sqrt(2)*1.96*Sw,
    CVw = 100*Sw/Xw
    )

df_bland96 %>%
  select(children, `1st`, `2nd`, `3rd`, `4th`) %>% mutate(children = factor(children)) %>%
  gather(key = measurement, value = value, -children) %>%
  # select only the eye ID, the eye ID, and the OCTA variable of interest
  aov(formula = value ~ children)
```

##Table 2: Repeatability

Table 2 shows repeatability estimates of vessel density measurements.

<span style="color:blue">This table was modeled after Table 2 in Venugopal (2018).[@venugopalRepeatabilityVesselDensity2018] They stratified these values by dichotomous glaucoma status, but we're not sure if that is appropriate.</span> 

```{r}

# function that gets repeatability statistics
get_repeatability <- function(df, octa_var){

  # make OCTA variable name
  octa_name <- var_labels[as.character(octa_var)][[2]]
octa_name
  # intra-visit dataset
  df %>%

    # select only the measurement ID and the OCTA variable of interest
    select(ptid, eyeid, eye_imaged, order_obs, !!octa_var) %>%

    # only select the first 2 measurements
    # filter(order_obs <= 2) %>%

    # group by measurement ID
    group_by(eyeid) %>%

    # calculate the mean and variance of measurements for each individual
    mutate(
      mean = base::mean(!!octa_var),
      var = stats::sd(!!octa_var)^2
      ) %>%

    # remove grouping by measurement ID
    ungroup() %>%

    # take the average of mean measurements and the average of the variances
    summarise(
      µ = mean(mean),
      Sw = sqrt(mean(var))
      ) %>%

    # square root the variances
    mutate(
      CRw = sqrt(2)*1.96*Sw,
      CVw = 100*Sw/µ,
      var_name = octa_name
      ) %>%
    select(var_name, everything())

  # octa_name
  # #Sw is the "residual standard error" from the anova output
  # data_octa_intra %>%
  #   # select only the eye ID, the eye ID, and the OCTA variable of interest
  #   select(Target, Judge, VAD) %>% mutate(Target = factor(Target)) %>%
  #   aov(formula = VAD ~ Target)

}

# call repeatability statistics on each measurement, then make into a table
get_repeatability(df = data_octa_global_intra, octa_var = quo(v_diameter)) %>%
  rbind(
    get_repeatability(df = data_octa_global_intra, octa_var = quo(v_area_density)),
    get_repeatability(df = data_octa_global_intra, octa_var = quo(v_skeleton_density)),
    get_repeatability(df = data_octa_global_intra, octa_var = quo(v_perimeter_index)),
    get_repeatability(df = data_octa_global_intra, octa_var = quo(v_complexity_index)),
    get_repeatability(df = data_octa_global_intra, octa_var = quo(flow_impair_zone)),
    get_repeatability(df = data_octa_global_intra, octa_var = quo(flux))
  ) %>%
  # make html "kable" and round digits
  kable(digits = 3) %>%
  kable_styling(
    #bootstrap table classes
    bootstrap_options = c( "hover")
  )
```

<span style="color:blue">I still need to figure out how they calculated the 95% CIs, which I think is just $1.96*\sqrt{SE}$ because they all appear symmetric, but I'm not sure yet. Then I need to stratify these stats by glaucoma diagnosis (yes/no). - DG 8/29/2019</span>

<span style="color:blue">Bruce showed me a paper that has the calculation of the confidence intervals. It's not difficult to manually code them, just need to read the paper. - DG 9/4/2019</span>

##Calculating Repeatability Statistics

Explanations of repeatability statistic calculations  

###Within-Subject Mean $(\mu_w)$

$\mu_w$ is the average measurement of both eyes for each individual. The overall mean $(\mu)$ is calculated by taking the mean of $\mu_w$ across the dataset. The overall mean is included in the final table for reference.

\[ \mu_w = \frac{1}{M}\sum_{i=1}^{M} x_i  \]

\[\text{where } M \text{ is the number of measurements per eye,}\]

###Within-Subject Standard Deviation $(S_w)$

Find $S_w$ by first calculating the variance of the measurements per individual eye. The equation below is general for any number of measurements, but in this study there are only 2 measurements per eye.  

\[ \sigma^2_{measurement} =  \frac{1}{M}\sum_{i=1}^{M} (x_i - \mu_w)^2 \]

\[ \text{where } M \text{ is the number of measurements per eye}  \]

To calculate $S_w$, average of the variance of measurements $(\sigma^2_{measurement})$ for all eyes measured, then take the square root.  

\[ Sw =  \sqrt{\frac{1}{N}\sum_{i=1}^{N}\sigma^2_{measurement,i}} \]

\[ \text{where } N \text{ is the number of eyes measured} \]

###Within-Subject Coefficient of Repeatability $(CR_w)$

The $CR_w$ provides the uncertainty of repeated measures.

\[ CRw = \sqrt2 * 1.96 * Sw \]

The $CR_w$ can be interpreted as:  

"The difference between two measurements for the same subject is expected to be less than [$CR_w$] for 95% of pairs of observations."[@blandStatisticsNotesMeasurement1996]

###Within-Subject Coefficient of Variation $(CV_w)$

\[ CVw = 100 *\frac{Sw}{\mu} \]


#Intraclass Correlation Coefficient

Intraclass Correlation Coefficient (ICC)

```{r}
# function that produces a kable output for each dataset
icc_tabulate <- function(df){

  library(ICC)

  # function that calls ICC function, adds a column to front with variable name
  icc_that <- function(df, var){
      c(var, ICCest(ptid, var, data = df))
  }

  # call the ICC function on each OCTA variable, then r bind into a table
  rbind(
    icc_that(df, "v_diameter"),
    icc_that(df, "v_area_density"),
    icc_that(df, "v_skeleton_density"),
    icc_that(df, "v_perimeter_index"),
    icc_that(df, "v_complexity_index"),
    icc_that(df, "flow_impair_zone"),
    icc_that(df, "flux")
  ) %>% as.data.frame() %>%
    # round columns
    mutate(
      ICC = as.double(ICC),
      LowerCI = as.double(LowerCI),
      UpperCI = as.double(UpperCI),
      k = as.double(k),
      varw = as.double(varw),
      vara = as.double(vara)
      ) %>%

    # call kable
    kable(digit = 3) %>%
    #bootstrap table classes
    kable_styling(bootstrap_options = c("striped", "hover"))

}
```

##Intra-Visit ICC

```{r, warning = FALSE}
icc_tabulate(data_octa_global_intra)
```

##Inter-Visit ICC

```{r, warning = FALSE}
icc_tabulate(data_octa_global_inter)
```

##Calculating ICC Statistic

\[ \text{ICC} = \frac{\text{between}}{\text{between} + \text{within}}\]

<span style="color:blue">"In the output for the random effects model ‘SID’ is the estimated between-group variance, which in the ICCest output is called ‘vara’ (variance among groups). The within-group variance is ‘Residual’ in the random effects model output, while in ICCest it’s called ‘varw.’ You can see that the values for those two variances match perfectly in the two different outputs. The ICC is defined as between/(between+within). That calculation is not part of the random effects model output, but it is part of the ICCest output, along with confidence intervals for the ICC estimate. The ICC are really high for intravisit (around 97%) and a little lower for intervisit (about 94%). If we were to calculate these ICC for glaucoma and controls separately I think they would be significantly lower because of the reduced between-group variability." - BB 9/4/2019</span> 

#Mixed Effects Regression Models


##Model Fitting

<span style="color:blue">"For the markdown file let’s just spit out all the output from the random effects model and the iccest command so we can show it to Grace. Eventually we can put the ICC in the table with the other repeatability measures you have already tabled but I don’t think we need it for the meeting. I also added a model to test whether SS influences reliability. The reference level is ‘10’ now, which is what we want, and you can see that SS is highly associated with VAD, meaning it affects reliability." - BB 9/4/2019</span>

```{r}
model1.fit <- lmer(v_area_density ~ (1|ptid), data=data_octa_global_intra)
summary(model1.fit)

# model2.fit <- lmer(v_area_density ~ SS + (1|ptid), data=data_octa_global_intra)
# summary(model2.fit)
```

#References